### 2020-01-10 ISSTA/implementation.tex
issta shrink start

The implementation of the above ideas is conceptually straightforward,
and should be easy to apply in other automated test generation
settings with a natural notion of actions.  We present our
implementation by first sketching a mapping from the formalism to one
particular popular way to represent automatically generated unit
tests, in Section \ref{sec:pools}.  The particular implementations of
horizontal (Section \ref{sec:horizimp}) and vertical (Section
\ref{sec:vertimp}) determinism checks, as well as probabilistic
delta-debugging support (Section \ref{sec:ddimp}) are described after a general
overview of the TSTL testing tool (Section \ref{sec:tstlintro}).

@@

\begin{enumerate}
\item A declaration of a set of array variables, referred to as
  \emph{value pools}, e.g., {\tt int [] intVP = new int[3];}.
  \item A set of assignments of constant values to elements of
    primitive type value pools, e.g. {\tt intVP[2] = 1;}.
    \item A part in which all statements are assignments of calls of a
      method to a value pool, with all arguments to the method also
      taken from a value pool, e.g. {\tt intVP[2] = fooObjectVP[2].bar(intVP[0],
        intVP[1]);} (these are referred to as \emph{array-canonical} statements).
    \end{enumerate}

@@

\subsection{A Brief Introduction to TSTL}
\label{sec:tstlintro}

We implemented our approach in the TSTL \cite{NFM15,ISSTA15,tstlsttt}
system, an open-source
language and tool for property-based testing
\cite{Hypothesis,ClaessenH00} of Python code.   TSTL, including full
support for our nondeterminism detection methods, is available at
\url{https://github.com/agroce/tstl}, and the {\tt examples} directory
contains the source code needed to duplicate all of our
experiments\footnote{In some cases, you will also need to install an
  appropriate version of the tested software, by doing, e.g., {\tt pip
    install pyfakefs}.};
TSTL can also be installed using Python's {\tt pip} tool, by typing
{\tt pip install tstl}; you will need to clone the github repository
to have access to the case studies used in the paper, however.

@@

\begin{figure}
  {\scriptsize
\begin{code}
pool: <int> 5
pool: <l> 5
pool: <s> 5
\vspace{0.1in}
<int> := <[1..20]>
<l> := []
<l> := list(<s>)
<s> := set()
<s> := set(<l>)
\vspace{0.1in}
<l>.append(<int>)
\{ValueError\} <l>.remove(<int>)
<s>.add(<int>)
\{KeyError\} <s>.remove(<int>)
\vspace{0.1in}
property: len(<l>) >= len(set(<l,1>))
property: len(<s>) == len(list(<s,1>))
\end{code}
}
\caption{Simple TSTL harness testing lists and sets}
\label{fig:simple}
\end{figure}

In TSTL, a test consists of a sequence of
\emph{actions}, where actions execute arbitrary Python code, but are
expected to work by modifying the \emph{state} of a test, which is
stored in a fixed set of pools containing Python objects.  TSTL's
approach essentially follows the pool-based canonical form defined by
Andrews et al. and described above, but relaxes many of its restrictions
(e.g., a constant value can be used in a method call without first
assigning it to a pool). The
TSTL actions defined by a test harness correspond to the set $A$ of
actions in Section \ref{sec:formal}, and transitions between states
result from executing out the Python code in an action; these actions
are, roughly, the array-canonical statements of the value pool formalism.  Figure
\ref{fig:simple} shows a very simple TSTL harness that tests Python's
built-in {\tt set} and {\tt list} types.  The harness defines the
(visible) state of a test as consisting of three named pools of Python
objects, {\tt int}, {\tt l}, and {\tt s}.  These pools have no
inherent associated type, though a type can be annotated as an
additional constraint that TSTL will check holds during test
execution.  Initially all the pools hold the special Python value {\tt
  None} indicating they have not been initialized.  In terms of our formalism,
$I = \{${\tt [<int> = [None, None, None, None, None]}, {\tt <l> =
  [None, None, None, None, None]}, {\tt <s> = [None, None, None, None,
  None]]}$\}$.  Actions with the {\tt :=} assignment operator
initialize pools.  For example, the line of code {\tt l := []} defines a set of 5
actions, each of which initializes a different member of the list pool
{\tt l} to an empty list --- {\tt l0 = []} ... {\tt l4 = []}.  Actions
without a {\tt :=} require that pools used in them already be
initialized.  TSTL provides constructs like {\tt <[1..20]>} for
choosing values in a range for use in a test, and allows a user to
specify which exceptions an action may raise without causing the test
to fail (as in {\tt \{KeyError\} <s>.remove(<int>)}).

@@

@@

\begin{table*}
\centering
{\scriptsize

@@

Note that TSTL provides explicit support for exploring the real-world sources of
nondeterminism described in Section \ref{sec:sources}, by supporting
execution in a new process ({\tt processNondeterministic} API call and
{\tt --checkProcessDeterminism} command line argument) and allowing
control over timing changes in re-execution, by specifying a delay to
introduce between actions (parameters to API calls, {\tt
  --determinismDelay} command line arugment).  Changing the timing of
steps is very effective when exact timing is somehow involved in
results, and often effective in exposing concurrency-related
nondeterminism, since the delay will frequently change the
interleaving of threads.

@@

\subsection{Tools for Nondeterministic Delta-Debugging}
\label{sec:ddimp}

@@

\subsection{Key Challenges}
The key challenge in all of these implementations was choosing
appropriate default values.  In particular, how many times to retry
tests (horizontal determinism) or actions (vertical/failure determinism), what delays to use, and numbers of replications
required in nondeterministic delta-debugging.  Because our confidence
that these defaults are ideal is low, we made it easy to change the
values used, as shown by the presence of command line options and
optional function parameters in Tables \ref{tab:methods} and \ref{tab:options}.

>>>>

\begin{comment}
The implementation of the above ideas is conceptually straightforward,
and should be easy to apply in other automated test generation
settings with a natural notion of actions.  We present our
implementation by first sketching a mapping from the formalism to one
particular popular way to represent automatically generated unit
tests, in Section \ref{sec:pools}.  The particular implementations of
horizontal (Section \ref{sec:horizimp}) and vertical (Section
\ref{sec:vertimp}) determinism checks, as well as probabilistic
delta-debugging support (Section \ref{sec:ddimp}) are described after a general
overview of the TSTL testing tool (Section \ref{sec:tstlintro}).
\end{comment}

@@

\begin{enumerate}
\item A declaration of a set of array variables, referred to as
  \emph{value pools}, e.g., {\tt int [] intVP = new int[3];}.
  \item A set of assignments of constant values to elements of
    primitive type value pools, e.g. {\tt intVP[2] = 1;}.
    \item A part in which all statements are assignments of calls of a
      method to a value pool, with all arguments to the method also
      taken from a value pool, e.g. {\tt intVP[2] = fooVP[2].bar(intVP[0],
        intVP[1]);} (these are referred to as \emph{array-canonical} statements).
    \end{enumerate}

@@

%\subsection{A Brief Introduction to TSTL}
%\label{sec:tstlintro}

We implemented our approach in the TSTL \cite{NFM15,ISSTA15,tstlsttt}
system, an open-source
language and tool for property-based testing
\cite{Hypothesis,ClaessenH00} of Python code.   TSTL, including full
support for our nondeterminism detection methods, is available at
\url{https://github.com/agroce/tstl}, and the {\tt examples} directory
contains the source code needed to duplicate all of our
experiments\footnote{In some cases, you will also need to install an
  appropriate version of the tested software, by doing, e.g., {\tt pip
    install pyfakefs}.}.
\begin{comment}
TSTL can also be installed using Python's {\tt pip} tool, by typing
{\tt pip install tstl}; you will need to clone the github repository
to have access to the case studies used in the paper, however.

@@

\begin{figure}
  {\scriptsize
\begin{code}
pool: <int> 5
pool: <l> 5
pool: <s> 5
\vspace{0.1in}
<int> := <[1..20]>
<l> := []
<l> := list(<s>)
<s> := set()
<s> := set(<l>)
\vspace{0.1in}
<l>.append(<int>)
\{ValueError\} <l>.remove(<int>)
<s>.add(<int>)
\{KeyError\} <s>.remove(<int>)
\vspace{0.1in}
property: len(<l>) >= len(set(<l,1>))
property: len(<s>) == len(list(<s,1>))
\end{code}
}
\caption{Simple TSTL harness testing lists and sets}
\label{fig:simple}
\end{figure}
\end{comment}

In TSTL, a test consists of a sequence of
\emph{actions}, where actions execute arbitrary Python code, but are
expected to work by modifying the \emph{state} of a test, which is
stored in a fixed set of pools containing Python objects.  TSTL's
approach essentially follows the pool-based canonical form defined by
Andrews et al. and described above, but relaxes many of its restrictions
(e.g., a constant value can be used in a method call without first
assigning it to a pool). The
TSTL actions defined by a test harness correspond to the set $A$ of
actions in Section \ref{sec:formal}, and transitions between states
result from executing out the Python code in an action; these actions
are, roughly, the array-canonical statements of the value pool
formalism.

\begin{comment}
  Figure
\ref{fig:simple} shows a very simple TSTL harness that tests Python's
built-in {\tt set} and {\tt list} types.  The harness defines the
(visible) state of a test as consisting of three named pools of Python
objects, {\tt int}, {\tt l}, and {\tt s}.  These pools have no
inherent associated type, though a type can be annotated as an
additional constraint that TSTL will check holds during test
execution.  Initially all the pools hold the special Python value {\tt
  None} indicating they have not been initialized.  In terms of our formalism,
$I = \{${\tt [<int> = [None, None, None, None, None]}, {\tt <l> =
  [None, None, None, None, None]}, {\tt <s> = [None, None, None, None,
  None]]}$\}$.  Actions with the {\tt :=} assignment operator
initialize pools.  For example, the line of code {\tt l := []} defines a set of 5
actions, each of which initializes a different member of the list pool
{\tt l} to an empty list --- {\tt l0 = []} ... {\tt l4 = []}.  Actions
without a {\tt :=} require that pools used in them already be
initialized.  TSTL provides constructs like {\tt <[1..20]>} for
choosing values in a range for use in a test, and allows a user to
specify which exceptions an action may raise without causing the test
to fail (as in {\tt \{KeyError\} <s>.remove(<int>)}).

@@

\end{comment}

@@

\begin{comment}
\begin{table*}
\centering
{\scriptsize

@@

\end{comment}

Note that TSTL provides explicit support for exploring the real-world sources of
nondeterminism described in Section \ref{sec:sources}, by supporting
execution in a new process and allowing
control over timing changes in re-execution, by specifying a delay to
introduce between actions.  Changing the timing of
steps is very effective when exact timing is somehow involved in
results, and often effective in exposing concurrency-related
nondeterminism, since the delay will frequently change the
interleaving of threads.

@@

\begin{comment}
\subsection{Tools for Nondeterministic Delta-Debugging}
\label{sec:ddimp}

@@

\subsection{Key Challenges}
The key challenge in all of these implementations was choosing
appropriate default values.  In particular, how many times to retry
tests (horizontal determinism) or actions (vertical/failure determinism), what delays to use, and numbers of replications
required in nondeterministic delta-debugging.  Because our confidence
that these defaults are ideal is low, we made it easy to change the
values used, as shown by the presence of command line options and
optional function parameters in Tables \ref{tab:methods} and
\ref{tab:options}.
\end{comment}
